---
title: "multivariate analysis"
author: "Rodrigue"
date: "2023-11-23"
output:
  html_document: default
---

- Set the mean vector and covariance matrix
The code defines a 2x2 covariance matrix with values of 2, 0.5, 0.5, and 1 as well as a mean vector with values of 1 and 2. Using the mvrnorm function from the MASS package, it creates 1000 multivariate normal random samples, providing the seed for reproducibility and utilizing a tryCatch block to manage problems.

```{r}
library(MASS)
mean_vector <- c(1,2)
cov_matrix <- matrix(c(2,0.5,0.5,1), nrow = 2)
random_samples <- mvrnorm(n = 1000, mu = mean_vector, Sigma = cov_matrix)
colnames(random_samples) <- c('iPhone','Techno')
head(random_samples)
mean_vector
cov_matrix
```
- set the generated samples as scatter plot
The first and second columns of the samples matrix are used by this R code to generate a scatter plot.
```{r}
plot(random_samples[,1:2], main = "Scatter Plot of Bivariate Normal Distribution",
     xlab = "Types of phone", ylab = "Demands/preferences")
```
The function generates scatterplot matrices using the R lattice module. While "splom(~random_samples[,1:2])" and "splom(~random_samples[,1:2], pch = 16, col = 'blue')" build scatter plot matrices with distinct plot features, such as point shape and color, the "pairs(random_samples)" function creates a matrix of scatterp lots.

```{r} 
library(lattice)
# let generate scatter plot by grouping
pairs(random_samples)
splom( ~ random_samples[,1:2])
```
- correlation matrix (by curiosity):
```{r}
correlation_matrix <- cor(random_samples)
corrplot::corrplot(correlation_matrix, method = "ellipse")
```
- Create a grid for a bivariate normal distribution.

```{r}
# Let create the grid of values
x <- seq(-2,4,length.out = 100)
y <- seq(-2,4,length.out = 100)
grid <- expand.grid(x = x,y = y)
grid
```
- Calculate the bivariate normal density for each point in the grid.
```{r}
library(dplyr)
library(MASS)
# The bivariate normal density for each point in the grid
density_values <- dmvnorm(grid, mean = mean_vector, sigma = cov_matrix)
rounded_density <- round(density_values, digits = 2)
# let match the grid dimensions
density_matrix <- matrix(density_values, nrow = length(x), ncol = length(y), byrow = TRUE)
# let have a look at the plot
filled.contour(x, y, density_matrix, main = "Grid for a Bivariate Normal Density",
               xlab = "sequences x-axes", ylab = "sequences y-axes", color.palette = terrain.colors)
```
- Plot the bivariate normal density using a contour plot, labeling the axes appropriately.
```{r}
# Plot the bivariate normal density using a contour plot
contour(x, y, density_matrix, main = "Contour plot of Bivariate Normal Density",
        xlab = "sequences x-axes", ylab = "sequences y-axes")
```

Calculate Probabilities:
- Choose a region of interest (e.g., X < 2 and Y < 3) and calculate the probability.
- Print the calculated probability.

This R function determines the likelihood of a bivariate normal distribution inside a given region of interest using the mvtnorm package. After defining the mean, covariance matrix, and region boundaries, pmvnorm calculates the probability inside the selected area.

```{r}
library(mvtnorm)
lower_limits <- c(-Inf,-Inf)
upper_limits <- c(2,3)
# the region of interest
probability <- pmvnorm(lower = lower_limits, upper = upper_limits,
                        mean = mean_vector, sigma = cov_matrix)
# the calculated probability
cat("Probability in the region of interest:", probability, "\n")
```

Test for Multivariate Normality:
- Take a sample of at least 100 observations and two variables.

A synthetic dataset that represents a bivariate normal distribution is produced by this R function. The dataset is created using the mvrnorm function after the seed for reproducibility is established, a sample size of 100 is specified, the mean vector and covariance matrix are defined, and so on. The created distribution's initial few rows are printed at the end.

```{r}
# let create data vectors 
set.seed(1234) 
sample_size <-100  
# the mean_vector is having the two variables required
sample_meanvector <- mean_vector                                    
sample_covariance_matrix <- cov_matrix 
sample_distribution <- mvrnorm(n = sample_size, 
                               mu = sample_meanvector,  
                               Sigma = sample_covariance_matrix) 
# print top of distribution 
head(sample_distribution)
```
- Test if the dataset follows a multivariate normal distribution.
- Print and interpret the test results.

The 'MVN' library is used in these R code samples to evaluate multivariate normality using the Henze-Zirkler test ('hz'). The test result applied to the'sample_distribution' is stored in the'result' variable. To gain an understanding of the multivariate normality of the data, the next code would look over and analyze the test results that were acquired.

The p-values acquired from the Henze-Zirkler test are used to interpret the'result'. A low p-value suggests that multivariate normality has been faulty. There is evidence to reject normality if p < 0.05. On the other hand, p > 0.05 indicates that there is no significant variance, hence confirming the multivariate normality assumption in the results.

```{r}
library(MVN)
# Test for multivariate normality
result <- mvn(sample_distribution, mvnTest = "hz")
# Test result and interpretation
result
```